---
title: "Reproducible Research: Peer Assessment 1"
date: September 2018
author: DJR
output:
  html_document: default
keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Load libraires
```{r libraries, message= FALSE}
        library(lattice)
        #library(dplyr)
        #library(ggplot2)
        #library(impute)
        #library(lubridate)
        #library(Hmisc)
```
## Loading and preprocessing the data
1. Load the data (i.e. `read.csv()`)  
``` {r load}
origdf = read.csv(unzip("activity.zip"))
```
2. Process/transform the data (if necessary) into a format suitable for your analysis  
No process/transform needed yet.  

## What is mean total number of steps taken per day?
For this part of the assignment, you can ignore the missing values in  
the dataset.  
  
1. Make a histogram of the total number of steps taken each day  
```{r plot1}
        #Create data frame with sum of steps for each day.
        bydaydf<-aggregate(steps~date,origdf,sum)

        histogram(bydaydf$steps,xlab="Total Daily Steps", ylab= "Day Count", type= "count", nint=9)
```
  
2. Calculate and report the **mean** and **median** total number of steps taken per day  
```{r meanmedian1}
        mean1=mean(bydaydf$steps)
        median1=median(bydaydf$steps)
```

 Mean is `r format(mean1,digits=2, nsmall=2)` steps per day.   
 Median is `r median1` steps per day.   
        

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
