---
title: "Reproducible Research: Peer Assessment 1"
date: September 2018
author: DJR
output:
  html_document: default
keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Load libraires
```{r libraries, message= FALSE}
library(lattice)
#library(dplyr)
#library(ggplot2)
#library(impute)
#library(lubridate)
#library(Hmisc)
```
## Loading and preprocessing the data
1. Load the data (i.e. `read.csv()`)  
``` {r load}
origdf <- read.csv(unzip("activity.zip"))
```
2. Process/transform the data (if necessary) into a format suitable for your analysis  
No process/transform needed yet.  

## What is mean total number of steps taken per day?
For this part of the assignment, you can ignore the missing values in  
the dataset.  
  
1. Make a histogram of the total number of steps taken each day  
```{r plot1}
#Create data frame with sum of steps for each day.
bydaydf <- aggregate(steps~date,origdf,sum)

histogram(bydaydf$steps,xlab="Total Daily Steps", ylab= "Day Count", type= "count", nint=9)
```
  
2. Calculate and report the **mean** and **median** total number of steps taken per day  
```{r meanmedian1}
mean1=mean(bydaydf$steps)
median1=median(bydaydf$steps)
```

 Mean is `r format(mean1,digits=2, nsmall=2)` steps per day.   
 Median is `r median1` steps per day.   
        

## What is the average daily activity pattern?
3. Make a time series plot (i.e. `type = "l"`) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)  
```{r plot2}
#Create data frame with average steps for each interval
byintervaldf <- aggregate(steps~interval,origdf,mean)
plot(byintervaldf$interval, byintervaldf$steps, type = "l", xlab = "5-minute interval of the day", ylab= "average steps")
```

4. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?  
```{r maxstepsinterval}
maxstepsint <- byintervaldf$interval[which.max(byintervaldf$steps)]
```

Interval number `r maxstepsint` contains the maximum number of steps accorss all days on average.  

## Imputing missing values




## Are there differences in activity patterns between weekdays and weekends?
